Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-04 paper/paper.tex
Added table of results
Fix up table
name of university
nudge intro

%\usepackage{makeidx}  % allows for indexgeneration
%
\title{Fully Automatic Loop-free Program Synthesis by Bounded Model Checking}
\author{Daniel Kroening \and Matt Lewis}
\institute{Oxford University}

@@

We demonstrate the effectiveness of our method by synthesizing several subtle
programs acting on machine integers, and a selection of floating point programs.
\end{abstract}

@@

Program synthesis is the mechanized construction of software that provably
satisfies a given specification.  Synthesis tools promise to relieve the
programmer from thinking about \emph{how} the problem is to be solved;
instead, the programmer only provides a compact description of \emph{what}
is to be achieved.  Foundational research in this area has been
exceptionally fruitful, beginning with Alonzo Church's work on the
\emph{Circuit Synthesis Problem} in the sixties~\cite{church-synth}.
In this paper, we present a new method that brings the promise of
automating programming closer to application.

\section{Related Work}
The closest work to ours is {\sc brahma}~\cite{brahma}, which is also a method for
synthesizing loop-free bitvector programs.  In~\cite{brahma} the authors characterise
the program synthesis problem as a second order $\exists \forall$ formula, and
described the CEGIS algorithm for solving these formulae.  This algorithm reduces
the satisfiability of the $\exists \forall$ formula to the satisfiability of several
first order formulae, each of which can be decided using an off-the-shelf SMT solver.
{\sc Brahma} takes as input a specifcation (an SMT formula) and a library of components
(one component is more or less a single machine instruction).  It then finds
a permutation of the component library which satisfies the specification.  This
permutation is the final synthesised program.  In the event that the library does
not contain enough components to synthesise a correct program, the user is told
that the library is too small and is required to add more components until
the synthesis succeeds.  Each round of synthesis amounts to solving an
$\exists \forall$ formula that is quadratic in the size of the component library.

Our work represents an incremental improvement on {\sc brahma}.  We require the
user to supply a program specification written in C, but then ask nothing more of her.
Our encoding allows us to generate an $\exists \forall$ formula which is \emph{linear}
in the length of the shortest program satisfying the user's specification,
which we then check for satisfiability using the CEGIS algorithm.  It should
be noted that since the programs produced by {\sc brahma} are permutations of the
component library, the shortest correct program can be no longer than the size
of the library.

Another successful approach to program synthesis has been program sketching, in
particular the {\sc sketch} tool~\cite{sketch}.  With this method the user supplies
a program sketch, written in a C-like language, together with either a specification
(written in the same language) or a suite of unit tests.  The sketch is a skeleton
program which contains much of the high level structure of the final program,
but with several ``holes'' that {\sc sketch} fills in to meet the specification,
again using the CEGIS algorithm.

@@

\section{Basic Synthesis Algorithm}

@@

>>>>

%\usepackage{makeidx}  % allows for indexgeneration
%
\title{Fully Automatic Loop-free Program Synthesis by Bounded Model Checking}
\author{Daniel Kroening \and Matt Lewis}
\institute{University of Oxford}

@@

We demonstrate the effectiveness of our method by synthesizing several subtle
programs acting on machine integers, and a selection of floating-point programs.
\end{abstract}

@@

Program synthesis is the mechanized construction of software that provably
satisfies a given specification.  Synthesis tools promise to relieve the
programmer from thinking about \emph{how} the problem is to be solved;
instead, the programmer only provides a compact description of \emph{what}
is to be achieved.  Foundational research in this area has been
exceptionally fruitful, beginning with Alonzo Church's work on the
\emph{Circuit Synthesis Problem} in the sixties~\cite{church-synth}.

In this paper, we present a new method that brings the promise of
automating programming closer to application.  We make pragmatic
restrictions of the general synthesis problem in order to arrive at an
effective solution.  To this end, we focus on the problem of synthesizing
loop-free programs over machine integers and floating-point number from a
specification given as program fragment.

Our work represents a substantial refinement of {\sc brahma}~\cite{brahma},
which is also the closest work to ours is.  {\sc brahma} is also a method
for synthesizing loop-free bitvector programs.  In~\cite{brahma}, the
authors characterise the program synthesis problem as a second-order
$\exists \forall$ formula, and introduce the CEGIS algorithm for solving
these formulae.  This algorithm reduces the satisfiability of the $\exists
\forall$ formula to the satisfiability of several first-order formulae, each
of which can be decided using an off-the-shelf SMT solver.  {\sc Brahma}
takes as input a specifcation (given as an SMT formula) and a library of
components (one component is more or less a single machine instruction).  It
then finds a permutation of the component library that satisfies the
specification.  This permutation is the final synthesised program.  In the
event that the library does not contain enough components to synthesise a
correct program, the user is told that the library is too small and is
required to add more components until the synthesis succeeds.  Each round of
synthesis amounts to solving an $\exists \forall$ formula that is quadratic
in the size of the component library.

We require the user to supply a program specification written in C, but then
ask nothing more of her.  Our encoding allows us to generate an $\exists
\forall$ formula that is \emph{linear} in the length of the shortest program
satisfying the user's specification, which we then check for satisfiability
using the CEGIS algorithm.  In contrast, the program generated by {\sc
brahma} can be no longer than the size of the library, as the synthesis
outcome is a permutation of the component library.

Another successful approach to program synthesis has been \emph{program
sketching}, in particular the {\sc sketch} tool~\cite{sketch}.  With this
method the user supplies a program sketch, written in a C-like language,
together with either a specification (written in the same language) or a
suite of unit tests.  The sketch is a skeleton program which contains much
of the high level structure of the final program, but with several ``holes''
that {\sc sketch} fills in to meet the specification, again using the CEGIS
algorithm.

@@

\section{The Basic Synthesis Algorithm}

@@

\bigskip

\begin{tabular}{l|l|l|l|l}
Problem & {\sc Brahma} & {\sc Kalashnikov} & {\sc Kalashnikov} & {\sc Kalashnikov} \\
        &               &                    & (no width)         & (CBMC only) \\
\hline
p1 & 3.20s &{\bf 1.83s} &23.54s &23.54s \\
p2 & 3.60s &{\bf 1.49s} &1271.12s &1271.12s \\
p3 & {\bf 1.40s} &1.60s &4.09s &4.09s \\
p4 & 3.30s &{\bf 1.86s} &79.72s &79.72s \\
p5 & 2.20s &{\bf 1.98s} &12.16s &12.16s \\
p6 & 2.40s &{\bf 1.51s} &104.19s &104.19s \\
p7 & {\bf 1.00s} &4.04s &- &- \\
p8 & {\bf 1.40s} &4.31s &- &- \\
p9 & {\bf 5.80s} &35.09s &505.10s &505.10s \\
p10 & 76.10s &{\bf 9.80s} &- &- \\
p11 & 57.10s &{\bf 9.40s} &{\bf 9.40s} &{\bf 9.40s} \\
p12 & 67.80s &{\bf 21.43s} &{\bf 21.43s} &{\bf 21.43s} \\
p13 & 59.60s &{\bf 27.74s} &{\bf 27.74s} &{\bf 27.74s} \\
p14 & 118.90s &{\bf 54.71s} &{\bf 54.71s} &{\bf 54.71s} \\
p15 & 62.30s &{\bf 43.49s} &{\bf 43.49s} &{\bf 43.49s} \\
p16 & 78.10s &{\bf 27.41s} &{\bf 27.41s} &{\bf 27.41s} \\
p17 & 45.90s &{\bf 27.73s} &{\bf 27.73s} &{\bf 27.73s} \\
p18 & 34.70s &{\bf 5.55s} &{\bf 5.55s} &{\bf 5.55s} \\
p19 & {\bf 108.40s} &- &- &- \\
p20 & {\bf 28.30s} &- &- &- \\
p21 & {\bf 279.00s} &- &- &- \\
p22 & {\bf 1668.00s} &- &- &- \\
p23 & {\bf 224.90s} &- &- &- \\
p24 & {\bf 2778.70s} &- &- &- \\
\end{tabular}
